---
title: "Assignment - 3"
author: "Matthew Farris"
date: "September 13, 2015"
output: html_document
---
## Wine Quality

As a Amatuer Wine connoisseur, I found the wine quality dataset to be the most intriquing, so I decided to pulled from it

```{r}
library(RCurl)



redwineURL <-getURL("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv")

redwine <-read.csv(text = redwineURL, sep =";")

whitewineURL <-getURL("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv")

whitewine <-read.csv(text = whitewineURL, sep =";")

summary(redwine)

summary(whitewine)
```

For this dataset, I thought it would be a good idea to find the best of the best. For my purpose here, I was extremely interested in the "rating". I did a bit of research and found that most sites use a similar criteria to this one (From Wine Spectactor).


|Rating | Descriptor | Description|  
|-------|------------|------------|  
|95-100 | Classic | a great wine|  
|90-94 | Outstanding | superior character and style|  
|80-89 | Good/Very Good | wine with special qualities|  
|70-79 | Average |  drinkable with minor flaws|  
|60-69 | Below average | drinkable but not recommended |  
|50-59 | Poor | undrinkable, not recommended|  


  
Furthermore, review of the website Wine Folly has found that ratings tend to follow a normal distribution. So, to check to see if this data follows this trend, I plotted a histogram of the data.  

```{r}
hist(redwine$quality, xlab = "Quality", main = "Red Wine Quality")

hist(whitewine$quality, xlab = "Quality", main = "White Wine Quality")
```

As the data appears to mimic the norm, I felt that, selecting all the great wines would be anything above a 7 on this scale, so this became the first criteria I wanted to use for my subset. 

Furthermore, I wanted to focus most on the areas that I believe most influence our precieved taste of a wine. When we think of wine tasting, we most often think of acidity, sweetness, and alcohol content. Along with the rating, this made 4 columns that I wished to focus on. 

```{r}

BestReds <- subset(redwine, quality >= 7, select = c(residual.sugar,pH,alcohol,quality))
summary(BestReds)

BestWhites <- subset(whitewine, quality >= 7, select = c(residual.sugar,pH,alcohol,quality))
summary(BestWhites)
```

So this is one way to sparse out the data that we have on hand. I felt that this was a good way to provide a summary of the 

I decided to take the subset one step further. In white wines, there is a number classification for the "sweetness" of the wine. It ranges from 00 to 10. The following shows how much sugar is consider "sweet or dry". 

|Sugar(g/l)*|Characteristic|
|-----|--------------|
| > 1 | Bone Dry|
| 1-5 | Dry|
|5 - 25| Off-Dry|
|25 - 45| Medium Dry|
|45 - 65| Sweet|
| < 65 | Very Sweet |

*Please note there is some variation in what is "Sweet" when it comes to sugarl levels. Some compare pH and sugar levels, which I may do at a later date, but as of now I just wanted to show some uses of subsets. I picked the most common sugar levels I could find. 

We see that none of the medium dry or sweet wines made it in the top white wines, so I decided to go back to the original data set, as a way to just visual some of the data using a box plot. I understand their are simplier ways to concatenate data like this (ie making a new column in the data frame and then creating a box plot), but I decided to go this route as proof of concept, and to highlight column name changes as this was necessary to produce a readable box plot. 



```{r}

Bonedry <- subset(whitewine, residual.sugar >= 0 & residual.sugar < 1, select= residual.sugar)
Dry <- subset(whitewine, residual.sugar >= 1  & residual.sugar < 5,  select= residual.sugar)
Off_Dry <- subset(whitewine, residual.sugar >= 5  & residual.sugar < 25,  select= residual.sugar)
Medium_Dry <- subset(whitewine, residual.sugar >= 25  & residual.sugar < 45,  select= residual.sugar)
Sweet <-subset(whitewine, residual.sugar >= 45  & residual.sugar < 65,  select= residual.sugar)
Very_Sweet <- subset(whitewine, residual.sugar >= 65,  select= residual.sugar)

names(Bonedry)[names(Bonedry) == "residual.sugar"] <- "Bonedry"
names(Dry)[names(Dry) == "residual.sugar"] <- "Dry"
names(Off_Dry)[names(Off_Dry) == "residual.sugar"] <- "Off_Dry"
names(Medium_Dry)[names(Medium_Dry) == "residual.sugar"] <- "Medium"
names(Sweet)[names(Sweet) == "residual.sugar"] <- "Sweet"
names(Very_Sweet)[names(Very_Sweet) == "residual.sugar"] <- "Very_Sweet"



boxplot(c(Bonedry,Dry,Off_Dry,Medium_Dry,Sweet,Very_Sweet), main = "Sweetness Distribution")
```


Citations:

https://vinewineandwander.wordpress.com/2012/03/10/0-10-decoded/

P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
Modeling wine preferences by data mining from physicochemical properties. In Decision Support Systems, Elsevier, 47(4):547-553, 2009.
